---
title: "AML HW 2"
author: "Shivangi Mundhra SUID:842548148"
date: '2022-09-19'
output: html_document
---

Read .csv file and store it into supermarketSales 

```{r}
supermarketSales <- read.csv("/Users/shivangi/Desktop/Applied Machine learning/Week 2/supermarket_sales.csv")
```

see the structure of the dataframe

```{r}
str(supermarketSales)
```


```{r}
library(tidyverse)
# install.packages("hrbrthemes")
library(hrbrthemes)
library(viridis)
library(forcats)
library(ggplot2)
histTotalSales <- supermarketSales %>% mutate(Branch = fct_reorder(Branch, Total)) %>% ggplot( aes(x = Total, color = Branch, fill = Branch)) + geom_histogram(alpha = 0.6, binwidth = 250) + theme(legend.position = "none", panel.spacing = unit(0.1, "lines"), strip.text.x = element_text(size = 8)) + xlab("Total Sales ($)") + ylab("Frequency") + facet_wrap(~Branch) + ggtitle("Histogram of Total Sales by Branch")
```

```{r}
histTotalSales
# + scale_fill_viridis(discrete = TRUE) + scale_color_viridis(discrete = TRUE) + theme_ipsum() 
```


```{r}
p1 <- ggplot(supermarketSales, aes(x = Branch, y = Unit.price, fill = Product.line)) + geom_boxplot() + facet_wrap(~Product.line) + ggtitle("Box plots of unit price by product line")
p1
```

create a chart that shows gross income by gender by branch

```{r}
# colored by the variable color
ggplot(supermarketSales, aes(x = Branch, y = gross.income)) + geom_bar(aes(color = Gender, fill = Gender), stat = "identity", position = position_stack()) + scale_color_manual(values = c("#0073C2FF", "#EFC000FF")) + scale_fill_manual(values = c("#0073C2FF", "#EFC000FF")) + ggtitle("Bar chart for gross income by gender by branch")
# Use position = position_dodge() 
p <- ggplot(supermarketSales, aes(x = Branch, y = gross.income)) + geom_bar(aes(color = Gender, fill = Gender), stat = "identity", position = position_dodge(0.8), width = 0.7) + scale_color_manual(values = c("#0073C2FF", "#EFC000FF"))+ scale_fill_manual(values = c("#0073C2FF", "#EFC000FF")) + ggtitle("Bar chart for gross income by gender by branch")
p
```

create a chart that shows number of sales by gender by branch

```{r}
byGenderSales <- supermarketSales %>% count(Branch, Gender, sort = TRUE)
barGS <- ggplot(data = byGenderSales, aes(x = Branch, y = n)) + geom_bar(aes(color = Gender, fill = Gender), stat = "identity", position = position_stack()) + ggtitle("Bar chart for number of sales by gender by branch")
barGS
# byGenderSales <- supermarketSales %>% count(Branch, Gender, sort = TRUE)
barGS1 <- ggplot(data = byGenderSales, aes(x = Branch, y = n)) + geom_bar(aes(color = Gender, fill = Gender), stat = "identity", position = position_dodge(0.8), width = 0.7) + ggtitle("Bar chart for number of sales by gender by branch")
barGS1
# position = position_dodge(0.8), width = 0.7
```

create a chart that shows number of sales by product type by branch

```{r}
byProductSales <- supermarketSales %>% count(Branch, Product.line, sort = TRUE)
barGS <- ggplot(data = byProductSales, aes(x = Branch, y = n)) + geom_bar(aes(color = Product.line, fill = Product.line), stat = "identity", position = position_stack()) + ggtitle("Number of sales by product line by branch")
barGS
# byGenderSales <- supermarketSales %>% count(Branch, Gender, sort = TRUE)
barGS1 <- ggplot(data = byProductSales, aes(x = Branch, y = n)) + geom_bar(aes(color = Product.line, fill = Product.line), stat = "identity", position = position_dodge(0.8), width = 0.7) + ggtitle("Number of sales by product line by branch")
barGS1
```

create a chart that shows time series of number of sales by branch

```{r}
supermarketSales$month <- months(as.Date(supermarketSales$Date))
```

```{r}
# supermarketSales[is.na(supermarketSales$month),]

supermarketSales$monthupdated <- supermarketSales$monthupdated
supermarketSales$monthupdated[substr(supermarketSales$Date,1,1) == '1'] <- 'January'
supermarketSales$monthupdated[substr(supermarketSales$Date,1,1) == '2'] <- 'February'
supermarketSales$monthupdated[substr(supermarketSales$Date,1,1) == '3'] <- 'March'
```

```{r}
byDateSales <- supermarketSales %>% count(Branch, monthNum = as.numeric(substr(supermarketSales$Date,1,1)), monthupdated, sort = TRUE)
byDateSales$monthupdated <- factor(byDateSales$monthupdated, levels = c("January", "February", "March"))
lineGS <- ggplot(data = byDateSales, aes(x = monthupdated, y = n, group = Branch, color = Branch)) + geom_line() + ggtitle("Number of sales by month by branch")
lineGS
# byGenderSales <- supermarketSales %>% count(Branch, Gender, sort = TRUE)
# barGS1 <- ggplot(data = byDateSales, aes(x = Branch, y = n)) + geom_bar(aes(color = Product.line, fill = Product.line), stat = "identity", position = position_dodge(0.8), width = 0.7)
# barGS1
byDateSales
```

create a chart that shows sales by times of day by branch

```{r}
supermarketSales$timeUpdated <- as.numeric(substr(supermarketSales$Time, 1, 2))
```

```{r}
byTimeSales <- supermarketSales %>% count(Branch, timeUpdated, sort = TRUE)
# byDateSales$monthupdated <- factor(byDateSales$monthupdated, levels = c("January", "February", "March"))
lineGS1 <- ggplot(data = byTimeSales, aes(x = timeUpdated, y = n, group = Branch, color = Branch)) + geom_line() + ggtitle("Number of sales by time of day by branch")
lineGS1
```

create a chart that shows gross income member type by branch

```{r}
byMemberRating <- supermarketSales %>% group_by(Branch, Customer.type) %>% summarize(meanRating = mean(Rating))
barGM <- ggplot(data = byMemberRating, aes(x = Branch, y = meanRating)) + geom_bar(aes(color = Customer.type, fill = Customer.type), stat = "identity", position = position_stack()) + ggtitle("Average rating by customer type and branch")
barGM
# byGenderSales <- supermarketSales %>% count(Branch, Gender, sort = TRUE)
barGM1 <- ggplot(data = byMemberRating, aes(x = Branch, y = meanRating)) + geom_bar(aes(color = Customer.type, fill = Customer.type), stat = "identity", position = position_dodge(0.8), width = 0.7) + ggtitle("Average rating by customer type and branch")
barGM1
```


create a chart that shows gross income by branch by product

```{r}
# colored by the variable color
ggplot(supermarketSales, aes(x = Branch, y = gross.income)) + geom_bar(aes(color = Product.line, fill = Product.line), stat = "identity", position = position_stack())
# + scale_color_manual(values = c("#0073C2FF", "#EFC000FF")) + scale_fill_manual(values = c("#0073C2FF", "#EFC000FF"))
# Use position = position_dodge() 
p <- ggplot(supermarketSales, aes(x = Branch, y = gross.income)) + geom_bar(aes(color = Product.line, fill = Product.line), stat = "identity", position = position_dodge(0.8), width = 0.7)
# + scale_color_manual(values = c("#0073C2FF", "#EFC000FF")) + scale_fill_manual(values = c("#0073C2FF", "#EFC000FF"))
p
```


create a chart that shows gross income by branch

```{r}
byPaymentType <- supermarketSales %>% count(Branch, Payment, sort = TRUE)
barGS <- ggplot(data = byPaymentType, aes(x = Branch, y = n)) + geom_bar(aes(color = Payment, fill = Payment), stat = "identity", position = position_stack()) + ggtitle("Number of sales by payment type by branch")
barGS
# byGenderSales <- supermarketSales %>% count(Branch, Gender, sort = TRUE)
barGS1 <- ggplot(data = byPaymentType, aes(x = Branch, y = n)) + geom_bar(aes(color = Payment, fill = Payment), stat = "identity", position = position_dodge(0.8), width = 0.7) + ggtitle("Number of sales by payment type by branch")
barGS1
```

